Continuous developmental change explains discontinuities in word learning

### 2019-08-27 journal/ms.tex
first version journal

\abstract{
``Cognitive development is often characterized in term of
discontinuities, but these discontinuities can sometimes be apparent
rather than actual and can arise from continuous developmental change.
To explore this idea, we use as a case study the finding by Stager and
Werker (1997) that children's early ability to distinguish similar
sounds does not automatically translate into word learning skills. Early
explanations proposed that children may not be able to encode subtle
phonetic contrasts when learning novel word meanings, thus suggesting a
discontinuous/stage-like pattern of development. However, later work has
revealed (e.g., through using more precise testing methods) that
children do encode such contrasts, thus favoring a continuous pattern of
development. Here we propose a probabilistic model describing how
development may proceed in a continuous fashion across the lifespan. It
characterizes developmental change as improvement in the precision of
perceptual encoding. Our model explained previous findings in the
literature and provided a new prediction --- the referents' visual
similarity modulates word leanring accuracy. The models' predictions
were corroborated by human data we collected from both preschool
children and adults. The braoder impact of this work is to show that
computational models, such as ours, can help us explore the extent to
which episodes of cogntive development that are typically thought of as
discontinuities may emerge from simpler, continuous mechanisms.''

@@

Nevertheless, at least in some cases, development may only appear to be
stage-like. Some discontinuities may be related to how we measure a
specific skill --- some tasks may not allow us to detect successful
learning. Other discontinuities may emerge due to statistical
thresholding (e.g., p-value \textless{} 0.05) which can create a
spurious dichotomy between success and failure in observing a given
behavior. In such cases, positing discontinuous stages is unnecessary.
Instead, a continuous model --- involving similar representations across
the lifespan --- may provide a simpler and more transparent account of
development.

We use a case study from word learning literature. Stager and Werker
(1997) first showed that children's early ability to distinguish similar
sounds does not automatically translate into word learning skills. The
authors measured word learning using an audio-visual habituation Switch
task. First, infants are familiarized with two word-object pairings
(e.g., lable A with object A and label B with object B). Second, they
are tested using two types of trials. The control \enquote{same} trial
consists of a correct pairing (e.g., label A with object A) and the
\enquote{switch} trial consists of a wrong pairing (e.g., label A with
object B). If babies have correctly learned the association during the
familiarization, they are supposed to be surprised by the
\enquote{switch} trial and not by the \enquote{same} trial. The former
should thus result in a greater looking time compared to the latter
(Werker, Cohen, Lloyd, Casasola, \& Stager, 1998).

@@

The initial discontinuous interpretation has been challenged by
subsequent work. For instance, Yoshida, Fennell, Swingley, and Werker
(2009) investigated whether failure in the switch task reflects a lack
of sound encoding during \emph{familiarization}, or whether it is only
due to the demands of the \emph{testing} method which does not allow
learning below a certain threshold to be detected. They used the same
familiarization procedure as Stager and Werker (1997), but instead of
comparing the looking times in \enquote{same} and \enquote{switch}
trials, they tested infants using a two-alternative choice task
comparing fixations to target and distractor objects (Fernald, Perfors,
\& Marchman, 2006; Golinkoff, Hirsh-Pasek, Cauley, \& Gordon, 1987).
Using this easier and more precise task, the researcher found evidence
for learning even in 14-month olds.

@@

Some researchers (Pajak et al., 2016; Swingley, 2007; Yoshida et al.,
2009) proposed that word knowledge may not be encoded in a binary
fashion, i.e., it is not the case that children either succeed or fail
in encoding minimal contrast when learning the meanings. Rather, they
may be encoding this knowledge in a graded fashion (ses also munakata,
2001). According to this view, development does not so much involve a
qualitative shift (i.e., a sudden emergence of an ability that did not
exist before) as much as it consists in the continuous refinement of
initially noisy knowledge.

@@

This general framework --- in which the precision of word knowledge is
characterize with the variance of a probability distribution --- can
already provide an intuitive way of thinking about several findings. In
particular, unlike the binary view, the probabilistic view allows for
the possibility that knowldege can be successful while still being
noisy. This new understanding can provide an account for the fact that
children show evidence of learning in some testing condition (Yoshida et
al., 2009), but not in others (Stager \& Werker, 1997).

In a word-pair learning paradigm, children are supposed to associate one
label, e.g, \enquote{bih}, with object 1 and a second label, e.g.,
\enquote{dih}, with object 2. Infants may succeed in learning both
associations. Nevertheless, the precision with which the words are
encoded can still be large, causing their probability distributions to
overalp (Figure \ref{fig:illus}, top). The way this knowledge is tested
can lead to divering results.

Failure in the Switch task (Stager \& Werker, 1997) can be exlained as
follows. Children are understood to succeed if they reject the wrong
pairing (e.g., \enquote{bih} with object 2). However, the overlap means
that children could think of \enquote{bih} as plausible misprononciation
of \enquote{dih}. They may require a stronger sound mismatch to
confidently reject the possibility that the wrong label is just an
acceptable deviation from the tyical pronouciation. In contrast, success
in the two-alternative choice task (Yoshida et al., 2009) occurs
precisely becasue children do not have to reject a wrong pairing.
Instead, the setting allows them to just choose which objects is a
better match to a given label, thus revealing whether they have a
prefernce, albeit samll, for the correct pairing.

In addition to explaining behavior in different tasks, the probabilsitic
understanding also accounts for different behaviors in the same task. In
particular, when the labels are quite distinct in the pereptual space
(\enquote{lif} vs. \enquote{neem}), the distributions do not overlap as
much as in the case of similar-sounding words (Figure \ref{fig:illus},
left). This fact means that the learners will have a stronger preference
fot the correct label-object pairings and less tolerance for the wrong
ones. That said, perceptual distinctiveness can be enhanced even for
minimally different sounds when other cues highlight their difference
(Dautriche, Swingley, \& Christophe, 2015; Rost \& McMurray, 2009, 2010;
Thiessen, 2007; Yeung \& Werker, 2009).

Finally, development can be understood as an increase in the precision
(i.e., a decrease in the variance) of the probabilistic representations,
leading to a lower overlap between the similar-sounding words (Figure
\ref{fig:illus}, right). Importantly, a more precise representation
still has a non-zero variance --- Learning difficulties can still be
induced with challenging stimuli or in cognitively demanding situations
as was demonstrate in adults studies (Pajak et al., 2016; White et al.,
2013).

@@

The present work proposes a model of word-pair learning based on the
probabilistic account. We tested the ability of this model to both
\emph{explain} various findings in previous experiments in both children
and adults (e.g., the fact that similar words are harder to learn than
different words) and to \emph{predict} new learning patterns that have
not been tested before (i.e., the effect of the referents' similarity on
word learning). The broader impact of this work is to explore the extent
to which development can be understood as a continuous refinement in
similar representations across the lifespan.

The paper is organized as follows. First, we introduce the model and we
explain how it allows us to characterize behavior in a word learning
task which resembles the one used in Stager and Werker (1997) and
Yoshida et al. (2009). Then we explore the predictions of the model
through simulating its behavior across different parameter settings.
Next we quantify the extent to which the model's predictions account for
human data we collected from both preschool children and adults.
Finally, we discuss the results in the lights of existing accounts of
word development.

@@

\caption{An illustration of the probabilistic/continuous account using simulated data. A word is represented with a distribution over the perceptual space (indicated in red or blue). When the uncertainty of the representation is large relative to the distance between the stimuli (top panel), an instance of the red category (indicated with a star) could also be a plausible instance of the green category, hence the low recognition accuracy score. The accuracy increases when the stimuli are less similar (left panel), or when the representation are more precise (right panel).}\label{fig:illus}
\end{figure}

@@

{\centering \includegraphics[width=300px]{figs/model}

@@

\[ p(o | C) \sim  \mathcal{N}(\mu_C, \sigma^2_C) \]

\[ p(s| L) \sim  \mathcal{N}(\mu_L, \sigma^2_L) \]

Finally, we assume there to be one-to-one mappings between concepts and
labels and that observers have successfully learned these mappings
during the exposure phase: \[
P(L_i|C_j) = 
\begin{cases}
  1 & \text{if  }  i=j \\  
  0  & \text{otherwise  }
\end{cases}
\]

@@

The learner hears a sound \(s\) and has to decide which object \(o\)
provides an optimal match to this sound (see Figure \ref{fig:task}). To
this end, they must compute the probability \(P(o|s)\) for all possible
objects. This probability can be computed by summing over all possible
concepts and labels:
\[P(o|s)=\sum_{C,L} P(o, C, L| s) \propto \sum_{C,L} P(o, C, L, s) \]

The joint probability \(P(o, C, L, s)\) is obtained by factoring the
Bayesian network in Figure \ref{fig:model}:
\[P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o) \]

which can be transformed using Bayes rule into:

\[P(o,C,L,s) = P(s|L)P(L|C)P(o|C)P(C) \]

Finally, assuming that the concepts' prior probability is uniformly
distributed\footnote{This is a reasonable assumption in our particular case given the similarity of the concepts used in each naming situation in our experiment.},
we obtain the following expression, where all conditional dependencies
are now well defined:

\[
P(o|s) = \frac{\sum_{C,L} P(s|L)P(o|C)P(L|C)}{\sum_{o} \sum_{C,L} P(s|L)P(o|C)P(L|C)}
\]

@@

We use the model to predict word learning in a task similar to the one
introduced by Stager and Werker (1997). We used a modified version of
the task where the testing method consists in a two-alternative forced
choice (Yoshida et al., 2009). In this task, participants are first
exposed to the association between pairs of nonsense words (e.g.,
\enquote{lif}/\enquote{neem}) and pairs of objects. The word-object
associations are introduced sequentially. After this exposure phase,
participants perform a series of test trials. In each of these trials,
one of the two sounds is uttered (e.g., \enquote{lif}) and participants
choose the corresponding object from the two alternatives. An overview
of the task is shown in Figure \ref{fig:task}.

From the general expression (1), we derive three exact analytical
solutions instantiating different learning assumptions. The first
solution is derived by assuming that the labels are recovered from
sounds with a certain level of uncertainty \(\sigma_L\), but that
concepts are unambiguously recovered from the observed objects, i.e.,
\(\sigma_C \rightarrow 0\). This assumption has been made --- whether
implicitly or explicitly --- by most previous work in this line of
research. One important implication of this assumption is that only the
similarity of word sounds modulates success in word learning, not the
similarity of the referents (as long as these referents are
differentiated perceptually). This assumption yields the following
probability function:

\begin{equation}
P(o_T|s)= \frac{1}{1 + e^{-\frac{\Delta s^2}{2\sigma_L^2}}}
\end{equation}

The second solution is derived by making the more general assumption
that both the labels and the concepts are recovered with ambiguity from
the sounds and objects. We first introduce the simplifying assumption
that the label-related uncertainty \(\sigma_L\) and the concept-related
uncertainty \(\sigma_C\) are of a similar magnitude, i.e.,
\(\sigma_C \approx \sigma_L = \sigma\). This assumption makes the
prediction that the sound similarity and the object similarity impact
word learning accuracy in exactly the same way. Furthermore, it allows
us to study the behavior of the model with minimal free parameters.

\begin{equation}
P(o_T|s)= \frac{1 + e^{- \frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}}{1 + e^{-\frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}+ e^{-\frac{\Delta s^2}{2\sigma^2}} + e^{-\frac{\Delta o^2}{2\sigma^2}}}
\end{equation}

@@

\begin{equation}
P(o_T|s)= \frac{1 + e^{- (\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}}{1 + e^{-(\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}+ e^{-\frac{\Delta s^2}{2\sigma_L^2}} + e^{-\frac{\Delta o^2}{2\sigma_C^2}}}
\end{equation}

In order to understand the predictions of the models (especially the
more general ones, i.e., Model 2 and 3), Figure \ref{fig:simulation}
show some simulations of the accuracy \(P(o_T|s)\) as a function of the
distinctiveness parameters (\(\Delta s\) and \(\Delta o\)) and the
uncertainty parameters \(\sigma_L\) and \(\sigma_C\).

\begin{figure}[htbp]
\centering
\includegraphics{ms_files/figure-latex/simulation-1.pdf}
\caption{\label{fig:simulation}The predicted probability of accurate
responses in the testing phase as a function of stimuli distinctiveness
\(\Delta s\) and \(\Delta o\) and representation precision \(\sigma\)
(For clarity, we assume here that \(\sigma\)=\(\sigma_C\)=\(\sigma_L\)).
Dashed line represents chance.}
\end{figure}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  For fixed values of \(\Delta o\) and \(\sigma\), the probability of
  accurate responses increases as a function of \(\Delta s\). This
  pattern accounts for the fact that similar sounds are generally more
  challenging to learn than different sounds for both children (Stager
  \& Werker, 1997) and adults (Pajak et al., 2016).
\item
  For fixed values of \(\Delta s\) and \(\Delta o\), accuracy increases
  when the representational uncertainty \(\sigma\) decreases. This fact
  may explain development, i.e., younger children have noisier
  representations (see Swingley, 2007; Yoshida et al., 2009), which
  leads to lower word recognition accuracy, especially for
  similar-sounding words.
\item
  For fixed values of \(\Delta s\) and \(\sigma\), accuracy increases
  with the visual distance between the semantic referents \(\Delta o\).
  This is a new prediction that our model makes. Previous work studied
  the effect of several bottom-up and top-down properties in
  disambiguating similar sounding words (e.g., Fennell \& Waxman, 2010;
  Rost \& McMurray, 2009; Thiessen, 2007), but to our knowledge, no
  previous study in the literature tested the effect of the visual
  distance between the semantic referents.
\end{enumerate}

@@

We report data from \(N=\) 63 children ages 4-5 years from the Bing
Nursery School on Stanford University's campus. An additional \(N=\) 39
children participated but were removed from analyses because they were
not above chance on the catch trials due to the challenging nature of
our procedure (see below). We also report data from \(N=\) 74 adult
participants tested on Amazon Mechanical Turk. An additional \(N=\) 26
were tested but removed from analyses because they had low scores on the
catch trials or because they were familiar with the non-English sound
stimuli we used in the adult experiment.

@@

As for the objects, we used the Dynamic Stimuli javascript
library\footnote{https://github.com/erindb/stimuli} which allowed us to
generate objects in four different categories: \enquote{tree,}
\enquote{bird,} \enquote{bug,} and \enquote{fish.} These categories are
supposed to be naturally occurring kinds that might be seen on an alien
planet. In each category, we generated \emph{different},
\emph{intermediate} and \emph{similar} pairs by manipulating a
continuous property controlling features of the category's shape (e.g,
body stretch or head fatness).

In order to validate and quantify our similarity scales, we ran a
separate survey on Amazon Mechanical Turk where we asked \(N=20\) adults
participants to evaluate the similarity of each sound and object pair on
a 7-point scale. Data are shown in Figure \ref{fig:stim} where we scaled
responses within the range {[}0,1{]} for each stimulus group. These data
will be used in all models as a proxy for the perceptual distance
between the sound pairs (\(\Delta s\)) and the object pairs
(\(\Delta o\)).

@@

Each age group saw only two of the three levels of similarity described
in the previous sub-section: \emph{different} vs. \emph{intermediate}
for the preschoolers, and \emph{intermediate} vs. \emph{similar} for
adults. We made this choice in light of pilot studies showing that
adults were at ceiling with \emph{different} sounds/objects, and
children were at chance with the \emph{similar} sounds/objects. That
said, this difference in the level of similarity is accounted for in the
model through using the appropriate distance (Figure \ref{fig:stim}).

@@

Preschoolers were tested at the nursery school using a tablet, whereas
adults used their own computers to complete the same experiment online.
Participants were tested in a sequence of five conditions: the four
experimental conditions plus the catch condition. In each condition,
participants saw a first block of four exposure trials followed by four
testing trials, and a second block of two exposure trials (for memory
refreshment) followed by an additional four testing trials. The length
of this procedure was demanding, especially for children, but we adopted
a fully within-subjects design based on pilot testing that indicated
that precision of measurement was critical for testing our experimental
predictions.

In the exposure trials, participants saw two objects associated with
their corresponding sounds. We presented the first object on the left
side of the tablet's screen simultaneously with the corresponding sound.
The second sound-object association followed on the other side of the
screen after 500ms. For both objects, visual stimuli were present for
the duration of the sound clip (about 800ms). In the testing trials,
participants saw both objects simultaneously and heard only one sound.
They completed the trial by selecting which of the two objects
corresponded to the sound. The object-sound pairings were randomized
across participants, as was the order of the conditions (except for the
catch condition which was always placed in the middle of the testing
sequence). We also randomized the on-screen position (left vs.~right) of
the two pictures on each testing trial.

@@

\fbox{\parbox[b][][c]{14cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/afourtassi/networks}}}
\vspace{1em}

>>>>

\abstract{
``Cognitive development is often characterized in terms of
discontinuities, but these discontinuities can sometimes be apparent
rather than actual and can arise from continuous developmental change.
To explore this idea, we use as a case study the finding by Stager and
Werker (1997) that children's early ability to distinguish similar
sounds does not automatically translate into word learning skills. Early
explanations proposed that children may not be able to encode subtle
phonetic contrasts when learning novel word meanings, thus suggesting a
discontinuous/stage-like pattern of development. However, later work has
revealed (e.g., through using more precise testing methods) that
children do encode such contrasts, thus favoring a continuous pattern of
development. Here we propose a probabilistic model that represents word
knowledge in a graded fashion and characterizes developmental change as
improvement in the precision of this graded knowledge. Our model
explained previous findings in the literature and provided a new
prediction --- the referents' visual similarity modulates word learning
accuracy. The models' predictions were corroborated by human data we
collected from both preschool children and adults. The broader impact of
this work is to show that computational models, such as ours, can help
us explore the extent to which episodes of cognitive development that
are typically thought of as discontinuities may emerge from simpler,
continuous mechanisms.''

@@

Nevertheless, at least in some cases, development may only appear to be
stage-like. Some discontinuities may be related to how we measure a
specific skill. Other discontinuities may emerge due to statistical
thresholding (e.g., p-value \textless{} 0.05) which can create a
spurious dichotomy between success and failure in observing a given
behavior. In such cases, positing discontinuous stages is unnecessary.
Instead, a continuous model --- involving similar representations across
the lifespan --- may provide a simpler and more transparent account of
development (see also McMurray (2007) and Shultz, Schmidt, Buckingham,
and Mareschal (1995)).

We use a case study from word learning literature. Stager and Werker
(1997) first showed that children's early ability to distinguish similar
sounds does not automatically translate into word learning skills. The
authors measured word learning using an audio-visual habituation Switch
task. First, infants are familiarized with two word-object pairings
(e.g., label 1 with object 1 and label 2 with object 2). Second, they
are tested using two types of trials. The control \enquote{same} trial
consists of a correct pairing (e.g., label 1 with object 1) and the
\enquote{switch} trial consists of a wrong pairing (e.g., label 1 with
object 2). If babies have correctly learned the association during the
familiarization, they are supposed to be surprised by the
\enquote{switch} trial and not by the \enquote{same} trial. The former
should thus result in a greater looking time compared to the latter
(Werker, Cohen, Lloyd, Casasola, \& Stager, 1998).

@@

The initial discontinuous interpretation has been challenged by
subsequent work. For instance, Yoshida, Fennell, Swingley, and Werker
(2009) investigated whether failure in the Switch task reflects a lack
of sound encoding during \emph{familiarization}, or whether it is only
due to the nature of the \emph{testing} method which does not allow
learning below a certain threshold to be detected. They used the same
familiarization procedure as Stager and Werker (1997), but instead of
comparing the looking times in \enquote{same} and \enquote{switch}
trials, they tested infants using a two-alternative choice task
comparing fixations to target and distractor objects (Fernald, Perfors,
\& Marchman, 2006; Golinkoff, Hirsh-Pasek, Cauley, \& Gordon, 1987).
Using this testing method, the researcher found evidence for learning
even in 14-month olds.

@@

Some researchers (Pajak et al., 2016; Swingley, 2007; Yoshida et al.,
2009) proposed that word knowledge may not be encoded in a binary
fashion, i.e., it is not the case that children either succeed or fail
in encoding minimal contrast when learning the meanings. Rather, they
may be encoding this knowledge in a graded fashion (see also Munakata
(2001)). According to this view, development does not so much involve a
qualitative shift (i.e., a sudden emergence of an ability that did not
exist before) as much as it consists in the continuous refinement of
initially noisy knowledge.

@@

This general framework --- in which the precision of word knowledge is
characterized with the variance of a probability distribution --- can
already provide an intuitive way of thinking about several findings. In
particular, unlike the binary view, the probabilistic view allows for
the possibility of word knowledge being both successful and noisy. This
new understanding can provide an account for the fact that children show
evidence of learning in some testing condition (e.g., Yoshida et al.,
2009) but not in others (e.g., Stager \& Werker, 1997) --- depending on
the precision of the measurement.

In a word-pair learning paradigm, children are supposed to associate one
label, e.g., \enquote{bih}, with object 1 and a second label, e.g.,
\enquote{dih}, with object 2. Infants may succeed in learning both
associations. Nevertheless, the variance with which the pair of words
are encoded can still be large, causing their probability distributions
to overlap (Figure \ref{fig:illus}, top). The way this (noisy) knowledge
is probed can lead to different results.

In the Switch task (Stager \& Werker, 1997), children are understood to
succeed if they reject a wrong association (e.g., \enquote{bih} with
object 2). However, a large overlap between \enquote{bin} and
\enquote{din} means that \enquote{bih} is itself a plausible
mispronunciation of \enquote{dih}. The wrong association may not be
rejected by children because the speaker could have said \enquote{bih}
but meant \enquote{dih}. In the two-alternative choice task (Yoshida et
al., 2009), children do not have to reject the wrong association; they
only need to show a preference for the correct one. Given a label, e.g.,
\enquote{bih}, they pick which object is a better match to this label.
Though it is possible that the speaker said \enquote{bih} and meant
\enquote{dih}, it is \emph{more} likely that the speaker both said and
meant \enquote{bih} --- this higher probability leads to a preference
for the correct object.

In addition to explaining the difference in behavior across the Switch
and the preferential looking tasks, the probabilistic account explains
difference in behavior within the same task. In particular, when the
labels are quite distinct in the perceptual space (\enquote{lif} vs.
\enquote{neem}), the probabilistic distributions do not overlap as much
as in the case of similar-sounding words (Figure \ref{fig:illus}, left).
This fact means that the learners will have less tolerance for the wrong
association, leading to a successful rejection in the Switch task (as
was reported by Stager and Werker (1997) and subsequent studies using
the same paradigm). Further, distinctiveness can be enhanced even for
minimally different sounds when other cues highlight their difference
(Dautriche, Swingley, \& Christophe, 2015; Rost \& McMurray, 2009, 2010;
Thiessen, 2007; Yeung \& Werker, 2009).

Finally, development can be understood as an increase in the precision
(i.e., a decrease in the variance) of the probabilistic knowledge,
leading to a lower overlap between the distributions of similar-sounding
words (Figure \ref{fig:illus}, right). Importantly, a more precise
representation still has a non-zero variance --- Learning difficulties
can still be induced with challenging stimuli or in cognitively
demanding situations as was demonstrated in adults studies (Pajak et
al., 2016; White et al., 2013).

@@

The present work proposes a model of word-pair learning based on the
probabilistic account. We tested the ability of this model to both
\emph{explain} various findings in previous experiments in both children
and adults (e.g., the fact that similar words are harder to learn than
different words) and to \emph{predict} new learning patterns that have
not been tested before (i.e., the effect of the referents' similarity on
word learning). We explore the extent to which the probabilistic account
allows us to understand development in terms of as a continuous
refinement in similar representations across the lifespan.

The paper is organized as follows. First, we introduce the model and we
explain how it allows us to characterize behavior in a
word-pair-learning paradigm. Then we explore the predictions of the
model through simulating its behavior across different parameter
settings. Next, we quantify the extent to which the model's predictions
account for human data we collected from both preschool children and
adults. Finally, we discuss the results in the lights of existing
accounts of word development.

@@

\caption{An illustration of the probabilistic/continuous account using simulated data. A word is represented with a distribution over the perceptual space (indicated in red or blue). When the uncertainty of the representation is large relative to the distance between the stimuli (top panel), an instance of the red category (indicated with a star) could also be a plausible instance of the green category, hence the low recognition accuracy score. The accuracy is higher when the stimuli are less similar (left panel), or when the representation are more precise (right panel).}\label{fig:illus}
\end{figure}

@@

{\centering \includegraphics[width=150px]{figs/model}

@@

\begin{equation} \label{eq:object}
 p(o | C) \sim  \mathcal{N}(\mu_C, \sigma^2_C) 
\end{equation}

\begin{center}
and 
\end{center}

\begin{equation} \label{eq:sound}
p(s| L) \sim  \mathcal{N}(\mu_L, \sigma^2_L)
\end{equation}

Finally, we assume there to be one-to-one mappings between concepts and
labels and that observers have successfully learned these mappings
during the exposure phase:

\begin{equation}
P(L_i|C_j) = 
\begin{cases}
  1 & \text{if  }  i=j \\  
  0  & \text{otherwise  }
\end{cases}
\end{equation}

@@

In our canonical inference case, the learner hears a sound \(s\) and has
to decide which object \(o\) provides an optimal match to this sound
(see Figure \ref{fig:task}). To this end, they must compute the
probability \(P(o|s)\) for all possible objects. This probability can be
computed by summing over all possible concepts and labels:

\begin{equation}
P(o|s)=\sum_{C,L} P(o, C, L| s)  
\end{equation}

Using the fact that \(P(o,C,L|s) = \frac{P(o,C,L,s)}{P(s)}\) and that
\(P(s)\) does not depend on \(o\), we arrive at the equation:

\begin{equation}
P(o|s) \propto \sum_{C,L} P(o, C, L, s) 
\end{equation}

The joint probability \(P(o,C,L,s)\) is obtained by factoring the
graphical model in Figure \ref{fig:model}:

\[P(o,C,L,s) = P(s|L)P(L|C)P(C|o)P(o)\]

Using Bayes' rule, we can rewrite \(P(C|o)\) in terms of \(P(o|C)\):

\[P(C|o) = \frac{P(o|C)P(C)}{P(o)}\]

By subtituting this term in the expression of the joint distribution
\(P(o,C,L,S)\) we obtain:

\[P(o,C,L,s) = P(s|L)P(L|C)P(o|C)P(C)\]

Finally, assuming that the concepts' prior probability \(P(C)\) is
uniformly
distributed,\footnote{This is a reasonable assumption in our particular case given the similarity of the concept pairs used in each naming situation in our experiment.}
we obtain the following expression, where all conditional dependencies
are now well defined:

\begin{equation} \label{eq:general}
P(o|s) \propto \sum_{C,L}  P(s|L)P(L|C)P(o|C)
\end{equation}

@@

We use the model to predict word learning in a task similar to the one
introduced by Stager and Werker (1997). We used a modified version of
the task where the testing method consists in a two-alternative
forced-choice (Yoshida et al., 2009). In this task, participants are
first exposed to two different word-object pairings (e.g., \enquote{lif}
- object 1, \enquote{neem} - object 2). The word-object associations are
introduced sequentially. After this exposure phase, participants perform
a series of test trials. In each of these trials, one of the two sounds
is uttered (e.g., \enquote{lif}) and participants choose the
corresponding object from the two alternatives. An overview of the task
is shown in Figure \ref{fig:task}.

\subsubsection{Model 1}\label{model-1}

From the general expression \ref{eq:general}, we derive three exact
analytical solutions instantiating different learning assumptions.
Recall from expressions \ref{eq:object} and \ref{eq:sound} that
\(P(o|C)\) and \(P(s|L)\) have parameters \(\sigma_C\) and \(\sigma_L\),
respectively, that control perceptual uncertainty. The first solution is
derived by assuming that the labels are recovered from sounds with a
certain level of uncertainty \(\sigma_L > 0\), but that concepts are
unambiguously recovered from the observed objects, i.e.,
\(\sigma_C \rightarrow 0\). This assumption has been made --- whether
implicitly or explicitly --- by most previous work in this line of
research. For example, in Stager and Werker (1997), the objects were
quite dissimilar. Thus, the assumption that they were easily
discriminated by infants seems relatively well justified. One important
implication of this assumption is that only the similarity of word
sounds modulates success in word learning, not the similarity of the
referents (as long as these referents are differentiated perceptually).
This assumption yields the following probability function:

\begin{equation} \label{eq:model1}
P(o_T|s)= \frac{1}{1 + e^{-\frac{\Delta s^2}{2\sigma_L^2}}}
\end{equation}

\subsubsection{Model 2}\label{model-2}

The second solution is derived by making the more general assumption
that both the labels and the concepts are recovered with noise from the
sounds and objects. We first introduce the simplifying assumption that
the label-related uncertainty \(\sigma_L\) and the concept-related
uncertainty \(\sigma_C\) are of a similar magnitude, i.e.,
\(\sigma_C \approx \sigma_L = \sigma\). This assumption makes the
prediction that the sound similarity and the object similarity impact
word learning accuracy in exactly the same way. Furthermore, it allows
us to study the behavior of the model with only one free parameter, an
important consideration given the small number of datapoints available
from any given infant experiment.

\begin{equation} \label{eq:model2}
P(o_T|s)= \frac{1 + e^{- \frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}}{1 + e^{-\frac{\Delta s^2 + \Delta o^2}{2\sigma^2}}+ e^{-\frac{\Delta s^2}{2\sigma^2}} + e^{-\frac{\Delta o^2}{2\sigma^2}}}
\end{equation}

\subsubsection{Model 3}\label{model-3}

@@

\begin{equation} \label{eq:model3}
P(o_T|s)= \frac{1 + e^{- (\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}}{1 + e^{-(\frac{\Delta s^2}{2\sigma_L^2}+ \frac{\Delta o^2}{2\sigma_C^2})}+ e^{-\frac{\Delta s^2}{2\sigma_L^2}} + e^{-\frac{\Delta o^2}{2\sigma_C^2}}}
\end{equation}

In order to understand the predictions of the models (especially the
more general ones, i.e., Model 2 and 3), Figure \ref{fig:simulation}
show simulations of the accuracy \(P(o_T|s)\) as a function of the
distinctiveness parameters (\(\Delta s\) and \(\Delta o\)) and the
uncertainty parameters \(\sigma_L\) and \(\sigma_C\).

\begin{figure}[htbp]
\centering
\includegraphics{ms_files/figure-latex/simulation-1.pdf}
\caption{\label{fig:simulation}The predicted probability of accurate
responses in the testing phase as a function of stimuli distinctiveness
\(\Delta s\) and \(\Delta o\) and representation precision \(\sigma\)
(For simplicity, we use model 2, which assumes that
\(\sigma\)=\(\sigma_C\)=\(\sigma_L\)). Dashed line represents chance.}
\end{figure}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  For fixed values of \(\Delta o\) and \(\sigma\), the probability of
  accurate responses increases as a function of \(\Delta s\). This
  pattern accounts for the fact that similar sounds are generally more
  challenging to learn than different sounds for both children (Stager
  \& Werker, 1997) and adults (Pajak et al., 2016).
\item
  For fixed values of \(\Delta s\) and \(\Delta o\), accuracy increases
  when the representational uncertainty \(\sigma\) decreases. This
  observation provides a simple model for developmental change. Younger
  children have noisier representations (see Swingley, 2007; Yoshida et
  al., 2009), which leads to lower word recognition accuracy, especially
  for similar-sounding words.
\item
  For fixed values of \(\Delta s\) and \(\sigma\), accuracy increases
  with the visual distance between the semantic referents \(\Delta o\).
  This is a new prediction that our model makes. Previous work studied
  the effect of several bottom-up and top-down properties in
  disambiguating similar sounding words (e.g., Fennell \& Waxman, 2010;
  Rost \& McMurray, 2009; Thiessen, 2007), but to our knowledge, no
  previous study in the literature tested the effect of the visual
  distance between the semantic referents.
\end{enumerate}

@@

We report data from \(N=\) 63 children ages 4-5 years from the Bing
Nursery School on Stanford University's campus. An additional \(N=\) 39
children participated but were removed from analyses (using
preregistered exclusion criteria) because they were not above chance on
the catch trials due to the challenging nature of our procedure (see
below). We also report data from \(N=\) 74 adult participants tested on
Amazon Mechanical Turk. An additional \(N=\) 26 were tested but removed
from analyses (again, using preregistered exclusion criteria) because
they had low scores on the catch trials or because they were familiar
with the non-English sound stimuli we used in the adult experiment.

@@

As for the objects, we used the Dynamic Stimuli javascript
library\footnote{https://github.com/erindb/stimuli} which allowed us to
generate objects in four different categories: \enquote{tree,}
\enquote{bird,} \enquote{bug,} and \enquote{fish.} These categories were
described to participants as naturally occurring kinds on an alien
planet. In each category, we generated \emph{different},
\emph{intermediate}, and \emph{similar} pairs by manipulating a
continuous property controlling features of the category's shape (e.g.,
body stretch or head fatness).

In order to validate and quantify our similarity scales, we ran a
separate survey on Amazon Mechanical Turk where we asked \(N=20\) adults
participants to evaluate the similarity of each sound and object pair on
a 7-point scale. Data are shown in Figure \ref{fig:stim} where we scaled
responses within the range {[}0,1{]} for each stimulus group. We used
these data in all models as an empirical measurement of the perceptual
distance between the sound pairs and the object pairs. The use of
empirical measurement allows us to eliminate \(\Delta s\) and
\(\Delta s\) as free parameters (see Frank and Goodman (2012) and Xu and
Tenenbaum (2007) for a similar strategy).

@@

Each age group saw only two of the three levels of similarity described
in the previous sub-section: \emph{different} vs. \emph{intermediate}
for the preschoolers, and \emph{intermediate} vs. \emph{similar} for
adults. We made this choice in light of pilot studies showing that
adults were at ceiling with \emph{different} sounds/objects, and
children were at chance with the \emph{similar} sounds/objects. That
said, this difference in the level of similarity is accounted for in the
model: We used empirical distance measurement to fill in the appropriate
values of \(\Delta s\) and \(\Delta o\) for each age group.

@@

Preschoolers were tested at the nursery school using a tablet, whereas
adults used their own computers to complete the same experiment online.
Participants were tested in a random sequence of five conditions: the
four experimental conditions plus the catch condition. In each
condition, participants saw a first block of four exposure trials
followed by four testing trials, and a second block of two exposure
trials (for memory refreshment) followed by an additional four testing
trials. The length of this procedure was demanding, especially for
children, but we adopted a fully within-subjects design based on pilot
testing that indicated that precision of measurement was critical for
testing our experimental predictions.

In the exposure trials, participants saw two objects associated with
their corresponding sounds. We presented the first object on the left
side of the tablet's screen simultaneously with the corresponding sound.
The second sound-object association followed on the other side of the
screen after 500ms. For both objects, visual stimuli were present for
the duration of the sound clip (about 800ms). In the testing trials,
participants saw both objects simultaneously and heard only one sound.
They completed the trial by selecting which of the two objects
corresponded to the sound. The object-sound pairings were randomized
across participants, as was the order of the conditions (except for the
catch condition which was always placed in the middle of the testing
sequence). We also randomized the on-screen position (left vs.~right) of
the two pictures on each testing
trial.\footnote{The experiment can be viewed online at https://tinyurl.com/word-pair-learning-experiment}

@@

\fbox{\parbox[b][][c]{14cm}{\centering All data and code for these analyses are available at\ \url{https://github.com/afourtassi/kidswitch}}}
\vspace{1em}

\section{Appendix}\label{appendix}

We show in Table 2 the output of the regression model we used to analyze
behavioral data in the experiment.
